---
title: "Parsing 2025-2026"
output: html_document
---

Packages:
```{r}
library(tidyverse)
library(readr)
library(stringr)
library(vroom)

#setting it up in the right directory
knitr::opts_knit$set(root.dir = "C:/Users/krian/OneDrive/Documentos/Bioinfo/C.Elegans/data")
```

C. Elegans worms were treated with Ethanol, Yucca, and Resveratrol solutions and their average movements were recorded using a Wiggleometer, the data was uploaded onto a csv. There are currently 19 cvs that were recorded over the course of several months. This files serves to parse the data.

###Explanation:
*Ethanol*: resveratrol and yucca is dissolved in ethanol - used as a control to see the effects
*Yucca*: ***Treatment Group*** An alleged superfood with antioxidant properties, seeing the extent of health benefits by testing it on worms (plant spanning 50 genuses native to north america)
*Resveratrol*: ***positive control*** phylotoalexin that has anti-inflammatory properties found in Yucca 

**NT = no treatment
ETOH = ethanol
RV = resveratrol
YS = Yucca

###Extracted data:
- Average activity counts per data intervals
- Standard Error (later/optional)

###Excluded points:
- first 3 hours
- M9_CTRL (media control) 
- All mutants; DAF16X, TJ356GFP, MR13, OWL13 (ONLY INCLUDE N2 and NT_YS)

**Code Outline**
Beginning
run all libraries
 turn all lines into raw data
 (start marker) grep "Group Activity: Average Activity Counts per Data Interval" *look for core phrase only, formatting with "<<>>" is not consistant among files*
 (end marker) grep ""Group Activity: Standard Error"
 only take lines from between these two markers *we only care about this one table*
 save into a variable END
 
START - clean the data some more
- filter rows to only keep samples with "N2" OR "NT_YS" in the name
- remove timepoints <= 180 mins
- pivot into long format
- print to file 
DONE
 
 
```{r}

setwd("C:/Users/krian/OneDrive/Documentos/Bioinfo/C.Elegans/data") 

extract_avg_activity <- function(file) {
  
  raw <- iconv(readLines(file, warn = FALSE, encoding = "latin1"),
               from = "latin1", to = "UTF-8")
  
  start <- grep("Group Activity: Average Activity Counts per Data Interval", raw)[1]
  end   <- grep("Group Activity: Standard Error", raw)[1]
  
  table_lines <- raw[(start + 1):(end - 1)]
  
  df <- readr::read_csv(
    paste(table_lines, collapse = "\n"),
    col_types = readr::cols(.default = "c")
  )
  
  # 1. Keep only N2 and NT_YS groups BEFORE pivoting
  df <- df %>% 
    filter(str_detect(.[[1]], "N2") | str_detect(.[[1]], "NT_YS"))
  
  # 2. Remove empty columns BEFORE pivoting
  #df <- df[, colSums(!is.na(df)) > 0]
  
  # 3. Remove timepoints <= 180 BEFORE pivoting
  #    (we detect numeric column names)
  numeric_cols <- suppressWarnings(as.numeric(names(df)[-1]))
  keep_cols <- which(numeric_cols > 180) + 1  # +1 to skip group column
  
  df <- df[, c(1, keep_cols)]
  
  # 4. Now pivot safely
  long <- df %>%
    rename(group = 1) %>%
    pivot_longer(
      cols = -group,
      names_to = "Time_min",
      values_to = "AvgActivity"
    ) %>%
    mutate(
      Time_min = as.numeric(Time_min),
      AvgActivity = as.numeric(AvgActivity)
    ) %>%
    filter(!is.na(Time_min) & !is.na(AvgActivity))
  
  return(long)
}
# Example:
clean_filtered <- extract_avg_activity("02_26_26_N2_OAP_AND_KID_NT_ETOH_RV_YS_NG_LS_BL_LL_AP.csv")

#Save in another file
write_csv(clean_filtered,
          "C:/Users/krian/OneDrive/Documentos/Bioinfo/C.Elegans/clean data/02_26_26_cleaned.csv")
```

